---
title: "Normalization of mass intentity by RC and QC"
date: "`r Sys.time()`"
output:
  bookdown::html_document2:
    toc: yes
    toc_float: true
    number_sections: true
    fig_width: 10
    fig_height: 7
    fig_caption: true
    theme: spacelab
    highlight: pygments
---
```{r echo=FALSE,message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library("R.utils")
library(cluster)
library(fpc)
library(mixOmics)
library(cowplot)
library(MetaboQC)

library(gg.gap)
```



```{r echo=FALSE, warning=FALSE}
ff <- getwd()
id <- unlist(strsplit(ff, "/", fixed = TRUE))
id <- id[length(id)]
```

<b>Job ID:</b> `r id`

<br><br><br><br>

# The overview of data
<br>
We show the data information and the basic description for all data resources.
<br><br>
## The data file list for our analysis
In this part, we show the file size for each file from uploading files.<br>
```{r echo=FALSE,message=FALSE, warning=FALSE}
dfs <- read.csv("dfs.csv")
colnames(dfs) <- c("File name", "Size(M)")
kable(dfs)
```

## Reference compound (RC) peaks intensity distribution
We use the boxplot to show the difference among all groups of mass data.
<br>
```{r echo=FALSE,message=FALSE, warning=FALSE, fig.align='center'}
rc_data <- read.csv("rc_data.csv")
colnames(rc_data) <- c("group", "Intensity")
p <- ggplot(data = rc_data)
p + stat_boxplot(geom = "errorbar", width = 0.6, aes(x = group, y = Intensity, group = group)) +
  geom_boxplot(aes(x = group, y = Intensity, group = group, fill = factor(group)), color = "black")+ theme_bw()+theme(legend.position = "none")
```

## The PCA plot for raw mass intensity data by grouping

```{r echo=FALSE,message=FALSE, warning=FALSE, fig.align='center'}
data_raw_QC <- read.csv("data_raw_pca_QC.csv")
data_raw_pca <- data_raw_QC[, -c(1:4)]
data_raw_pca <- apply(data_raw_pca, 2, as.numeric)
pc <- pca(t(data_raw_pca), ncomp = 2)
groupname <- data_raw_QC[, "pici"]

pp <- data.frame(pc$loading)
test <- data.frame(pp, group = as.character(groupname))
g1 <- ggplot(data = test, (aes(test[, 1], test[, 2], fill = group, colour = group))) +
  geom_point(size = 2) +
  theme_bw() +
  guides(colour = guide_legend(override.aes = list(size = 2))) +
  stat_ellipse(geom = "polygon", alpha = 0.2) +
  labs(x = paste("PC1(", sprintf("%.1f", pc$sde[1]^2 / sum(pc$sde^2) * 100), "%)", spe = ""), y = paste("PC2(", sprintf("%.1f", pc$sde[2]^2 / sum(pc$sde^2) * 100), "%)", spe = "")) +
  theme(legend.position = "bottom")

g2 <- ggplot(data = test, (aes(test[, 1], test[, 2], fill = group, colour = group))) +
  geom_point(size = 2) +
  theme_bw() +
  guides(colour = guide_legend(override.aes = list(size = 2))) +
  labs(x = paste("PC1(", sprintf("%.1f", pc$sde[1]^2 / sum(pc$sde^2) * 100), "%)", spe = ""), y = paste("PC2(", sprintf("%.1f", pc$sde[2]^2 / sum(pc$sde^2) * 100), "%)", spe = "")) +
  theme(legend.position = "bottom")
plot_grid(g1, g2, nrow = 1, labels = LETTERS[1:2], align = c("v", "h"))
```

# Adjusting mass data by below methods
<br>

## Adjusted by RC
```{r echo=FALSE,message=FALSE, warning=FALSE, fig.align='center'}
data_raw_QC <- read.csv("data_normalize_QC.csv")
data_raw_pca <- data_raw_QC[, -c(1:4)]
data_raw_pca <- apply(data_raw_pca, 2, as.numeric)
pc <- pca(t(data_raw_pca), ncomp = 2)
groupname <- data_raw_QC[, "pici"]

pp <- data.frame(pc$loading)
test <- data.frame(pp, group = as.character(groupname))
g1 <- ggplot(data = test, (aes(test[, 1], test[, 2], fill = group, colour = group))) +
  geom_point(size = 2) +
  theme_bw() +
  guides(colour = guide_legend(override.aes = list(size = 2))) +
  stat_ellipse(geom = "polygon", alpha = 0.2) +
  labs(x = paste("PC1(", sprintf("%.1f", pc$sde[1]^2 / sum(pc$sde^2) * 100), "%)", spe = ""), y = paste("PC2(", sprintf("%.1f", pc$sde[2]^2 / sum(pc$sde^2) * 100), "%)", spe = "")) +
  theme(legend.position = "bottom")

g2 <- ggplot(data = test, (aes(test[, 1], test[, 2], fill = group, colour = group))) +
  geom_point(size = 2) +
  theme_bw() +
  guides(colour = guide_legend(override.aes = list(size = 2))) +
  labs(x = paste("PC1(", sprintf("%.1f", pc$sde[1]^2 / sum(pc$sde^2) * 100), "%)", spe = ""), y = paste("PC2(", sprintf("%.1f", pc$sde[2]^2 / sum(pc$sde^2) * 100), "%)", spe = "")) +
  theme(legend.position = "bottom")
plot_grid(g1, g2, nrow = 1, labels = LETTERS[1:2], align = c("v", "h"))
```



## Adjusted by QC based on z-score


```{r echo=FALSE,message=FALSE, warning=FALSE, fig.align='center'}
data_raw_QC <- read.csv("ms_QC.csv")
data_raw_pca <- data_raw_QC[, -c(1:4)]
data_raw_pca <- apply(data_raw_pca, 2, as.numeric)
pc <- pca(t(data_raw_pca), ncomp = 2)
groupname <- data_raw_QC[, "pici"]

pp <- data.frame(pc$loading)
test <- data.frame(pp, group = as.character(groupname))
g1 <- ggplot(data = test, (aes(test[, 1], test[, 2], fill = group, colour = group))) +
  geom_point(size = 2) +
  theme_bw() +
  guides(colour = guide_legend(override.aes = list(size = 2))) +
  stat_ellipse(geom = "polygon", alpha = 0.2) +
  labs(x = paste("PC1(", sprintf("%.1f", pc$sde[1]^2 / sum(pc$sde^2) * 100), "%)", spe = ""), y = paste("PC2(", sprintf("%.1f", pc$sde[2]^2 / sum(pc$sde^2) * 100), "%)", spe = "")) +
  theme(legend.position = "bottom")

g2 <- ggplot(data = test, (aes(test[, 1], test[, 2], fill = group, colour = group))) +
  geom_point(size = 2) +
  theme_bw() +
  guides(colour = guide_legend(override.aes = list(size = 2))) +
  labs(x = paste("PC1(", sprintf("%.1f", pc$sde[1]^2 / sum(pc$sde^2) * 100), "%)", spe = ""), y = paste("PC2(", sprintf("%.1f", pc$sde[2]^2 / sum(pc$sde^2) * 100), "%)", spe = "")) +
  theme(legend.position = "bottom")
plot_grid(g1, g2, nrow = 1, labels = LETTERS[1:2], align = c("v", "h"))
```

## Adjusted by QC based on LOESS
```{r echo=FALSE,message=FALSE, warning=FALSE, fig.align='center'}
data_raw_QC <- read.csv("data_loess_normalize_QC.csv")
data_raw_pca <- data_raw_QC[, -c(1:4)]
data_raw_pca <- apply(data_raw_pca, 2, as.numeric)
pc <- pca(t(data_raw_pca), ncomp = 2)
groupname <- data_raw_QC[, "pici"]

pp <- data.frame(pc$loading)
test <- data.frame(pp, group = as.character(groupname))
g1 <- ggplot(data = test, (aes(test[, 1], test[, 2], fill = group, colour = group))) +
  geom_point(size = 2) +
  theme_bw() +
  guides(colour = guide_legend(override.aes = list(size = 2))) +
  stat_ellipse(geom = "polygon", alpha = 0.2) +
  labs(x = paste("PC1(", sprintf("%.1f", pc$sde[1]^2 / sum(pc$sde^2) * 100), "%)", spe = ""), y = paste("PC2(", sprintf("%.1f", pc$sde[2]^2 / sum(pc$sde^2) * 100), "%)", spe = "")) +
  theme(legend.position = "bottom")

g2 <- ggplot(data = test, (aes(test[, 1], test[, 2], fill = group, colour = group))) +
  geom_point(size = 2) +
  theme_bw() +
  guides(colour = guide_legend(override.aes = list(size = 2))) +
  labs(x = paste("PC1(", sprintf("%.1f", pc$sde[1]^2 / sum(pc$sde^2) * 100), "%)", spe = ""), y = paste("PC2(", sprintf("%.1f", pc$sde[2]^2 / sum(pc$sde^2) * 100), "%)", spe = "")) +
  theme(legend.position = "bottom")
plot_grid(g1, g2, nrow = 1, labels = LETTERS[1:2], align = c("v", "h"))
```

## Adjusted by QC based on Single-Poly3
```{r echo=FALSE,message=FALSE, warning=FALSE, fig.align='center'}
data_raw_QC <- read.csv("data_spoly_normalize_QC.csv")
data_raw_pca <- data_raw_QC[, -c(1:4)]
data_raw_pca <- apply(data_raw_pca, 2, as.numeric)
pc <- pca(t(data_raw_pca), ncomp = 2)
groupname <- data_raw_QC[, "pici"]

pp <- data.frame(pc$loading)
test <- data.frame(pp, group = as.character(groupname))
g1 <- ggplot(data = test, (aes(test[, 1], test[, 2], fill = group, colour = group))) +
  geom_point(size = 2) +
  theme_bw() +
  guides(colour = guide_legend(override.aes = list(size = 2))) +
  stat_ellipse(geom = "polygon", alpha = 0.2) +
  labs(x = paste("PC1(", sprintf("%.1f", pc$sde[1]^2 / sum(pc$sde^2) * 100), "%)", spe = ""), y = paste("PC2(", sprintf("%.1f", pc$sde[2]^2 / sum(pc$sde^2) * 100), "%)", spe = "")) +
  theme(legend.position = "bottom")

g2 <- ggplot(data = test, (aes(test[, 1], test[, 2], fill = group, colour = group))) +
  geom_point(size = 2) +
  theme_bw() +
  guides(colour = guide_legend(override.aes = list(size = 2))) +
  labs(x = paste("PC1(", sprintf("%.1f", pc$sde[1]^2 / sum(pc$sde^2) * 100), "%)", spe = ""), y = paste("PC2(", sprintf("%.1f", pc$sde[2]^2 / sum(pc$sde^2) * 100), "%)", spe = "")) +
  theme(legend.position = "bottom")
plot_grid(g1, g2, nrow = 1, labels = LETTERS[1:2], align = c("v", "h"))
```


## Adjusted by QC based on Single-Poly4
```{r echo=FALSE,message=FALSE, warning=FALSE, fig.align='center'}
data_raw_QC <- read.csv("data_spoly4_normalize_QC.csv")
data_raw_pca <- data_raw_QC[, -c(1:4)]
data_raw_pca <- apply(data_raw_pca, 2, as.numeric)
pc <- pca(t(data_raw_pca), ncomp = 2)
groupname <- data_raw_QC[, "pici"]

pp <- data.frame(pc$loading)
test <- data.frame(pp, group = as.character(groupname))
g1 <- ggplot(data = test, (aes(test[, 1], test[, 2], fill = group, colour = group))) +
  geom_point(size = 2) +
  theme_bw() +
  guides(colour = guide_legend(override.aes = list(size = 2))) +
  stat_ellipse(geom = "polygon", alpha = 0.2) +
  labs(x = paste("PC1(", sprintf("%.1f", pc$sde[1]^2 / sum(pc$sde^2) * 100), "%)", spe = ""), y = paste("PC2(", sprintf("%.1f", pc$sde[2]^2 / sum(pc$sde^2) * 100), "%)", spe = "")) +
  theme(legend.position = "bottom")

g2 <- ggplot(data = test, (aes(test[, 1], test[, 2], fill = group, colour = group))) +
  geom_point(size = 2) +
  theme_bw() +
  guides(colour = guide_legend(override.aes = list(size = 2))) +
  labs(x = paste("PC1(", sprintf("%.1f", pc$sde[1]^2 / sum(pc$sde^2) * 100), "%)", spe = ""), y = paste("PC2(", sprintf("%.1f", pc$sde[2]^2 / sum(pc$sde^2) * 100), "%)", spe = "")) +
  theme(legend.position = "bottom")
plot_grid(g1, g2, nrow = 1, labels = LETTERS[1:2], align = c("v", "h"))
```


# The comparation of all methods
```{r echo=FALSE,message=FALSE, warning=FALSE, fig.align='center'}
mean <- read.csv("mean.csv")
mean <- data.frame(mean)
ss <- quantile(as.numeric(mean[, 2]),na.rm = TRUE)
g1 <- ggplot(mean, aes(x = id, y = as.numeric(intensity), color = group, group = group)) +
  geom_line() +
  labs(
    y = "Mean of peaks intensity", x = ""
  ) +
  theme(axis.text = element_blank()) +
  ylim(c(0, ss[5]))+theme_bw()


sd <- read.csv("sd.csv")
sd <- data.frame(sd)
ss <- quantile(as.numeric(sd[, 2]),na.rm = TRUE)
g2 <- ggplot(sd, aes(x = id, y = as.numeric(intensity), color = group, group = group)) +
  geom_line() +
  labs(y = "SD of peaks intensity", x = "") +
  theme(axis.text = element_blank()) +
  ylim(c(0, ss[5]))+theme_bw()


plot_grid(g1, g2, nrow = 2, labels = LETTERS[1:2], align = c("v", "h"))
```

<br>
<br><br><br><br>
