---
title: "Process 5 (Spoly4)"
date: "`r Sys.time()`"
output:
  bookdown::html_document2:
    toc: yes
    toc_float: true
    number_sections: true
    fig_width: 10
    fig_height: 7
    fig_caption: true
    theme: spacelab
    highlight: pygments
---

```{r echo=FALSE,warning=FALSE, message=FALSE}
library(ggrepel)
library(ggplot2)
library(gridExtra)
library(venn)
library(ggpolypath)
library(clusterProfiler)
```

```{r echo=FALSE, warning=FALSE}
ff <- getwd()
id <- unlist(strsplit(ff, "/", fixed = TRUE))
id <- id[length(id)]
```

<b>Job ID:</b> `r id`

<br><br><br><br>

# The PCA plot for peak data by grouping

```{r echo=FALSE,message=FALSE, warning=FALSE, fig.align='center'}
data_raw_QC <- read.csv("sample_mass_RC_QC_normalize_spoly4.csv")
data_raw_pca <- data_raw_QC[, -c(1:4)]
data_raw_pca <- apply(data_raw_pca, 2, as.numeric)
pc <- pca(t(data_raw_pca), ncomp = 2)
groupname <- data_raw_QC[, "group"]

pp <- data.frame(pc$loading)
test <- data.frame(pp, group = as.character(groupname))
g1 <- ggplot(data = test, (aes(test[, 1], test[, 2], fill = group, colour = group))) +
  geom_point(size = 2) +
  theme_bw() +
  guides(colour = guide_legend(override.aes = list(size = 2))) +
  stat_ellipse(geom = "polygon", alpha = 0.2) +
  labs(x = paste("PC1(", sprintf("%.1f", pc$sde[1]^2 / sum(pc$sde^2) * 100), "%)", spe = ""), y = paste("PC2(", sprintf("%.1f", pc$sde[2]^2 / sum(pc$sde^2) * 100), "%)", spe = "")) +
  theme(legend.position = "bottom")

g2 <- ggplot(data = test, (aes(test[, 1], test[, 2], fill = group, colour = group))) +
  geom_point(size = 2) +
  theme_bw() +
  guides(colour = guide_legend(override.aes = list(size = 2))) +
  labs(x = paste("PC1(", sprintf("%.1f", pc$sde[1]^2 / sum(pc$sde^2) * 100), "%)", spe = ""), y = paste("PC2(", sprintf("%.1f", pc$sde[2]^2 / sum(pc$sde^2) * 100), "%)", spe = "")) +
  theme(legend.position = "bottom")
plot_grid(g1, g2, nrow = 1, labels = LETTERS[1:2], align = c("v", "h"))
```


# Differential metabolites analysis
In this section, we use the wilcox rank test and OPLS-DA to analyze the mass intensity. All differential peaks conform to the threshold  {FDR<0.05 and FC (FC>2 or FC<0.5)} and VIP>1

## Differential nalysis by wilcox rank test

```{r  echo=FALSE, fig.align='center', warning=FALSE,split=TRUE, fig.height=7,fig.width=10}
file <- list.files()
dd <- grep("wilcox_all_5.csv", file, fixed = TRUE)
file_wilcox <- file[dd]
pi <- 1
gplot <- list()
for (f in file_wilcox)
{
  data <- read.csv(file = f)
  data <- data.frame(data)
  nn <- unlist(strsplit(f, "++", fixed = TRUE))
  ggr <- nn[1]
  data$logFC <- log2(abs(as.numeric(data[, 2])))
  data$lgFDR <- -log10(abs(data[, 4]))
  data$status <- rep("unchange", length(data$lgFDR))
  thres1 <- which(as.numeric(data[, "FDR"]) < 0.05)
  thres2 <- which(abs(as.numeric(data[, "FC"])) > 2 | abs(as.numeric(data[, "FC"])) < 0.5)
  thres <- intersect(thres1, thres2)
  data$status[thres] <- "change"
  p <- ggplot(data, aes(logFC, lgFDR)) +
    geom_point(aes(color = status), size = 1) +

    # geom_jitter(aes(color = significant),alpha=0.4)+

    # 点添加标签
    geom_text_repel(aes(logFC, lgFDR, label = ifelse((status == "change"), as.character(ID), ""), color = status),
      size = 3, show.legend
      = TRUE
    ) +
    # 坐标轴名称
    labs(x = expression(log[2](FC)), y = expression(-log[10](FDR)), title = ggr) +
    # log10(0.5)~1.3
    geom_hline(yintercept = 1.3, linetype = 4) +
    # log2(2)=1,log2(0.5)=-1
    geom_vline(xintercept = c(-1, 1), linetype = 4) +
    ylim(min(data$lgFDR), max(data$lgFDR))+theme_bw()

  gplot[[pi]] <- p
  pi <- pi + 1
}
names(gplot) <- paste0("p", c(1:length(gplot)))
spn <- ceiling(length(gplot) / 2)
plot_grid(plotlist = gplot, nrow = spn, labels = "AUTO")
```


## Analysis by OPLS-DA

```{r echo=FALSE, fig.align='center', warning=FALSE,split=TRUE, fig.height=7,fig.width=10}
file <- list.files()

dd <- grep("oplsda_data_5.csv", file, fixed = TRUE)
if (length(dd) == 0) {
  errm <- "*************Error : No model was built because the predictive component was not significant*************."
} else {
  errm <- ""
  file_popls <- file[dd]
  pi <- 1
  gplot <- list()
  for (f in file_popls)
  {
    data <- read.csv(file = f)
    nn <- unlist(strsplit(f, "++", fixed = TRUE))
    ggr <- nn[1]
    summary <- read.csv(paste0(ggr, "++oplsda_summary_5.csv"))
    g <- ggplot(data = data, (aes(x, y, fill = group, colour = group))) +
      geom_point(size = 2) +
      theme_bw() +
      guides(colour = guide_legend(override.aes = list(size = 2))) +
      stat_ellipse(geom = "polygon", alpha = 0.2) +
      labs(x = "O1", y = "P1", title = ggr) +
      theme(legend.position = "right") +
      annotation_custom(grob = tableGrob(summary, rows = NULL))
    gplot[[pi]] <- g
    pi <- pi + 1
  }
  names(gplot) <- paste0("p", c(1:length(gplot)))
  spn <- ceiling(length(gplot) / 2)
  plot_grid(plotlist = gplot, nrow = spn, labels = "AUTO")
}
```


<b>`r errm`</b>


# The differrntial metabolites Venn diagram

```{r echo=FALSE, fig.align='center', warning=FALSE,split=TRUE, fig.height=7,fig.width=10}
file <- list.files()

dd <- grep("VIP_mean_5.csv", file, fixed = TRUE)
if (length(dd) == 0) {
  errm <- "*************Error : No Data for VIP>1*************"
} else {
  errm <- ""
  file_popls <- file[dd]
  pi <- 1
  gplot <- list()
  for (f in file_popls)
  {
    data <- read.csv(file = f)
    nn <- unlist(strsplit(f, "++", fixed = TRUE))
    ggr <- nn[1]
    wilcox <- read.csv(paste0(ggr, "++wilcox_mean_5.csv"))
    wilcox <- as.character(wilcox[, 1])
    oplsda <- as.character(data[, 1])
    da <- list(wilcox = wilcox, oplsda = oplsda)
    g <- venn(da, zcolor = "style", borders = FALSE, ggplot = TRUE, par = FALSE)
    g <- g + annotate("text", x = 270, y = 34, label = ggr)
    gplot[[pi]] <- g
    pi <- pi + 1
  }
  names(gplot) <- paste0("p", c(1:length(gplot)))
  spn <- ceiling(length(gplot) / 2)
  plot_grid(plotlist = gplot, nrow = spn, labels = "AUTO")
}
```

<b>`r errm`</b>


# THe metabolites matching in public database


## The information for differential metabolites (Top 50)
```{r echo=FALSE}
file <- list.files()

dd <- grep("DMC_wilcox_opls_mean_5.csv", file, fixed = TRUE)
if (length(dd) > 0) {
  peaks_mean <- read.csv("DMC_wilcox_opls_mean_5.csv", header = TRUE)
  kable(peaks_mean[1:50, 1:5])
}
```


## The identification of differential metabolites
This section uses the refMet of our database,including AIICCs, MassBank, KEGG, ChEBI, HMDB and NPAtlas, to identify the differential metabolites.
```{r echo=FALSE,fig.align='center', warning=FALSE,split=TRUE, message=FALSE, fig.height=7,fig.width=10}

file <- list.files()

dd <- grep("DMC_wilcox_opls_mean_identify_5.csv", file, fixed = TRUE)
if (length(dd) > 0) {
  errm <- ""
  peaks_mean <- read.csv("DMC_wilcox_opls_mean_identify_5.csv", header = TRUE)
  res <- peaks_mean %>%
    group_by(extend.DB) %>%
    summarise(n = length(unique(CID)))
  res <- data.frame(res)
  ggplot(data = res, aes(x = extend.DB, y = n, fill = extend.DB), fill = extend.DB) +
    geom_bar(stat = "identity") +
    geom_text(mapping = aes(x = extend.DB, y = n, label = n), size = 3)+theme_bw()+
    theme(legend.position = "none") +
    coord_flip()
} else {
  errm <- "*************No data for our results !***********"
}
```
<a href="DMC_wilcox_opls_mean_identify_5.csv">Download data of all database</a>


<b>`r errm`</b>

## The related pathways of differential metabolites
```{r echo=FALSE,fig.align='center', warning=FALSE,split=TRUE, message=FALSE, fig.height=10,fig.width=15}
wpgmtfile <- "../../refMet/hmdb_compound_pathway.gmt"
wp2gene <- read.gmt(wpgmtfile)
colnames(wp2gene) <- c("PID", "gene")
pathway <- read.csv("../../refMet/human_pathway_pathbank.csv")
pathway <- data.frame(pathway[, c(1, 2)])
colnames(pathway) <- c("PID", "name")
genelist <- read.csv("DMC_wilcox_opls_mean_identify_5.csv")
group <- as.character(genelist[, 2])
ugg <- unique(group)
m <- length(ugg)
errm <- rep("", m)
names(errm) <- ugg
gplot <- list()
for (i in 1:m) {
  sg <- which(group == ugg[i])
  gene <- as.character(genelist[sg, "extend1.ID"])
  gene <- gene[!is.na(gene)]
  result <- enricher(gene, TERM2GENE = wp2gene, TERM2NAME = pathway)
  if (length(result)  != 2) {
    errm[i] <- "--> No gene can be mapped....\n
--> return NULL..."
  } else {
    errm[i] <- ""
    pp <- paste0(ugg[i], "++DMC_wilcox_opls_mean_identify_pathway_5.csv")
    write.csv(result, pp)
    # barplot(ewp, showCategory=20)
    # cnetplot(ewp, foldChange=gene, circular = TRUE, colorEdge = TRUE)
    # edox <- setReadable(ewp, 'org.Hs.eg.db', 'ENTREZID')
    g <- cnetplot(result, foldChange = gene, categorySize = "p.adjust", colorEdge = TRUE)
    gplot[[i]] <- g
  }
}
if (length(gplot) > 0) {
  spn <- ceiling(length(gplot) / 2)
  plot_grid(plotlist = gplot, nrow = spn, labels = "AUTO")
}
```
<b>`r errm`</b>


## The related target proteins of differential metabolites {.tabset}

### Target protein (Top 20)
```{r echo=FALSE,warning=FALSE}
file <- list.files()
dd <- grep("DMC_wilcox_opls_mean_identify_target_5.csv", file, fixed = TRUE)
if (length(dd) > 0) {
  errm <- ""
  data <- read.csv("DMC_wilcox_opls_mean_identify_target_5.csv")
  dat <- data[1:20, c("name", "Group", "general.name", "gene.name", "gene.function", "unport.ID", "genecard.ID")]
  kable(dat)
} else {
  errm <- "There is no data for results!"
}
```


### Enzyme (Top 20)
```{r echo=FALSE,warning=FALSE}
file <- list.files()
dd <- grep("DMC_wilcox_opls_mean_identify_enzyme_5.csv", file, fixed = TRUE)
if (length(dd) > 0) {
  errm <- ""
  data <- read.csv("DMC_wilcox_opls_mean_identify_enzyme_5.csv")
  dat <- data[1:20, c("name.x", "Group", "general.name", "ECID", "name.y")]
  colnames(dat) <- c("name", "Group", "general.name", "ECID", "name")
  kable(dat)
} else {
  errm <- "There is no data for results!"
}
```

## {-}
<b>`r errm`</b>
<a href="DMC_wilcox_opls_mean_identify_target_5.csv">Download data of target protein</a><br>
<b>`r errm`</b>
<a href="DMC_wilcox_opls_mean_identify_enzyme_5.csv">Download data of enzyme protein</a>

<br><br><br><br>
